---
title: "GC.Size.Genes.Analysis"
author: "Irecha"
date: '2022-10-01'
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = "C:/Users/alam_/OneDrive/Documentos/R/Fusarium_Project")
```
## Statistical Analysis

This script is to perform statistical analysis to basic genome features to see if there is any correlation between them or any difference related to pathogenicity in several Fusarium species and forma speciales of Fusarium oxysporum.

### Load the necessary libraries using pacman:
```{r, warning = FALSE, message = FALSE}
pacman::p_load(RColorBrewer, ggplot2, dplyr, tidyverse, ggpubr)
```

### Load the metadata file
```{r, warning = FALSE, message = FALSE}
Fusarium_metadata <- read.csv(file = "Data/Fusarium_info.csv")
rownames(Fusarium_metadata) <- c(1:nrow(Fusarium_metadata))
Fusarium_metadata$Size <- Fusarium_metadata$Size/1000000 #Convert size variable to Mb units.
```

A positive correlation between the genomic features Genome Size and GC Content is often seen among Fusarium species and it's thus used as a genetic marker. In order to choose which correlation analysis to perform, it is necessary to determine if the variables follow a normal distribution:

```{r, warning = FALSE, message = FALSE}
shapiro.test(Fusarium_metadata$GC_Content)
shapiro.test(Fusarium_metadata$Size)

```

As seen with the Shapiro Test, variables don't follow a normal distribution. 
Plotting the variables also confirms this observation:

```{r, warning = FALSE, message = FALSE}
model <- lm(Size..bp. ~ GC.content...., data = Fusarium_metadata)
plot(model)
plot(Fusarium_metadata$Size)
plot(Fusarium_metadata$GC_Content)

```

Therefore, a Spearman correlation test is appropiated.

```{r, warning = FALSE, message = FALSE}
summary(model)
shapiro.test(resid(model))
Corr_test <- cor.test(Fusarium_metadata$Size..bp., Fusarium_metadata$GC.content...., method = "spearman", conf.level = 0.95, exact = FALSE)
Corr_test

```

Now, we can make a plot that shows the results of the Spearman correlation test:

```{r, warning = FALSE, message =  FALSE}
Plot1 <- ggplot(data = Fusarium_metadata, mapping = aes(Size, GC_Content)) + 
  geom_point(aes(colour = Pathogenicity), show.legend = TRUE, size = 3) + theme_minimal() +
  ylab('GC %') + xlab('Genome Size (Mb)') + geom_smooth(method=lm, level = 0.90, color = "red") +
  ggtitle("Correlation between G-C content and Genome Size among Fusarium oxysporum formae speciales and vanilla-associated Fusarium") +
  theme(plot.title = element_text(size = 14, face = "bold"))
Plot_correlation <- Plot1 + annotate("text", label = paste("Spearman R = ", as.character(round(Corr_test$estimate, digits = 2)), "p-value = ", as.character(round(Corr_test$p.value, digits = 2))), x=65, y=41, size = 5)
Plot_correlation
ggsave(filename = "Plots/GC_Size_correlation.png", plot = Plot_correlation, device = "png", bg=NULL, width = 1900, height = 1280, units = "px")

```

Next, we have to look for differences among the means of each genomic feature grouping the data by pathogenicity. The Wilcoxon-Matt-Whitney method for mean differences is the appropriated method given the non-normality condition of the data.
GC Content plot:

```{r, warning = FALSE, message = FALSE}
Plot2 <- ggplot(data = Fusarium_metadata, mapping = aes(Pathogenicity, GC_Content, fill = Pathogenicity, colour = Pathogenicity)) +
  geom_boxplot(alpha = 0.15, show.legend = TRUE, width = 0.25) + theme_classic() + ylab('GC %') + xlab(NULL) +
  stat_compare_means(method = "wilcox.test", label.x = 1, label.y = 42)
Plot2

```

Genome Size plot:

```{r, warning = FALSE, message = FALSE}
Plot3 <- ggplot(data = Fusarium_metadata, mapping = aes(Pathogenicity, Size, fill = Pathogenicity, colour = Pathogenicity)) +
  geom_boxplot(alpha = 0.15, show.legend = TRUE, width = 0.25) + theme_classic() + ylab('Genome Size (Mb)') + xlab(NULL) +
  stat_compare_means(method = "wilcox.test", label.x = 1, label.y = 65)
Plot3

```

Gene Content plot:

```{r, warning = FALSE, message = FALSE}
Plot4 <- ggplot(data = Fusarium_metadata, mapping = aes(Pathogenicity, Gene.content, fill = Pathogenicity, colour = Pathogenicity)) +
  geom_boxplot(alpha = 0.15, show.legend = TRUE, width = 0.25) + theme_classic() + ylab('Gene content') + xlab(NULL) +
  stat_compare_means(method = "wilcox.test", label.x = 1, label.y = 19000)
Plot4

```

Finally, we can group the three means difference plots into a single plot, to facilitate its interpretation:

```{r, warning = FALSE, message = FALSE}
Plot5 <- ggarrange(Plot2, Plot3, Plot4, nrow = 3, common.legend = TRUE, legend = "right", widths = 0.5)
Plot5
Plot_final <- annotate_figure(Plot5, top = text_grob("GC Content, Genome Size and Gene Content by Pathogenicity", 
                                      color = "black", face = "bold", size = 14))
Plot_final

ggsave("PDF/GC_Size_Gene_differences.pdf", plot = Plot_final, device = "pdf", units = "px", width = 2900, height = 1960)

```
